Introductory Physics Labs: A Tale of Two Transformations

### 2021-03-30 manuscript/rev1docs/response.tex

\begin{document}
\begin{letter}{Dr.\ Gary White \\ Editor: \textit{The Physics Teacher} \\ American Association
    of Physics Teachers \\ College Park, MD}

  \opening{Dear Dr.\ White and referee,}

  Thank you for your thoughtful review of our article entitled \textit{Introductory Physics
    Labs: A Tale of Two Transformations}. We have considered your review and are responding to
  your feedback in the following ways.

  \begin{description}
    \item[Institutional support for students needs] Dr.\ White wanted to know more about the
    ways that our institution attempted to support students during the pandemic.  We discuss
    academic measures taken by the university such as extending the withdrawal date, exempting
    withdrawals during the Spring 2020 and Fall 2020 semesters from the overall limit on
    withdrawals, and allowing students to be graded pass/fail after seeing their final grade
    (and having that grade not count toward their GPA).
    \item[Comparing ADI to Modeling Instruction] We've added a few sentences to the
    introduction discussing Modeling Instruction.  We also discuss ISLE (Investigative Science
    Learning Environment) as it is also a similar pedagogical approach driven by learning
    cycles.  Both Modeling and ISLE tend to be implemented in active/interactive lecture
    courses, while ADI was designed to mimic the ways that scientists learn in the laboratory,
    and has been designed to be implemented in laboratory courses regardless of the approach
    used in lecture. We find it to be particularly useful in our context as students in each of
    our lab courses take two different intro physics lecture courses (algebra-based and
    calculus-based).
    \item[Clarity about proposals and proposal approvals] We've added a few sentences talking
    about how we generated datasets during Spring 2020 and approve proposals in general. We
    want to point out that we do allow some proposals to go forward even if they have a flaw
    once data is collected.  Students are free to revise their proposals and gather more
    data. Dealing with unexpected results can be a powerful teacher. As far as creating
    datasets goes, we tried to make these data as raw as possible, for example sharing videos
    of a object oscillating on a spring and the mass of that object rather than giving them
    mass and period data directly. 
    \item[Details about Fall 2020] For the Fall 2020 semester, all data were collected by the
    students.  This includes images and videos as appropriate, often collected with students'
    smartphones. Additional text has been added to Section \textrm{IV}.B.
    \item[Details about the survey] The survey was conducted during the spring/early summer of
    2020.  Results were released to the faculty before the Fall 2020 semester began.  This was
    intended to help the faculty better understand the difficulties that were impacting our
    students. The date of the survey has been added to the text in the conclusion.
  \end{description}

### 2021-03-30 manuscript/onlineLabPaper.tex
Manuscript updated, word doc generated, response to editor and referees crafted.  Will submit to TPT in the morning.

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and
techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging students
in and assessing scientific practices, we secured funding, both internally and from the NSF, to
change this \emph{status quo} in physics, biology, and chemistry.  All three departments
transformed their labs using the Argument Driven Inquiry
(ADI)\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has been
specifically designed to engage students in authentic science practices. ADI is very much in
the spirit of transformations such as Modeling Physics\cite{modelingBrewe2008} and the
Investigative Science Learning Environment (ISLE) \cite{etkina2007investigative}. Oftentimes,
both Modeling Physics and ISLE are implemented in studio-style or even lecture classrooms,
whereas ADI is lab-focused. The common use of hybrid courses for Modeling and ISLE imply a
shorter time scale for their learning cycles, for example, an ISLE cycle often takes place over
the course of a week, whereas the cycles that we use in ADI take place over 4 weeks. For our
institutional transformation of our physics labs with ADI, we also utilized the AAPT Lab
Recommendations to inform our curricular choices.\citep{kozminski2014aapt}

@@

We developed lab kits with supplies that allowed students to perform the investigations outside
the teaching laboratory so we could provide students with a hands-on experience. We purchased
the lab kit items in collaboration with our campus bookstore, and the students purchased the
lab kits from the bookstore. We ordered lab kit items in bulk and, where possible, directly from
manufacturers to reduce the costs of the items. Each Physics 1 kit cost \$20.60 and each
Physics 2 kit cost \$39.00. Students payed for these kits in place of a lab manual, which we
provided online for this course. Lab kit materials were used for each investigation. The two video 
analysis investigations in the Physics 1 laboratory (one comparing the accelerations of a marble 
rolling up and down an incline and the other analyzing a marble collision in two dimensions) 
required the students to record their own videos of marbles from their lab kits using their 
smartphone cameras. The image analysis investigation in the 
Physics 2 laboratory required students to cut their own slits in aluminum foil and also create mounts 
for hairs to produce diffraction patterns with lab-kit lasers. The online adaptation of the face-to-face 
ADI curriculum has changed slightly. Pre-labs are done in groups asynchronously.  Proposals are 
developed in groups during the lab period, and scheduled argumentation sessions occur via 
WebEx during the lab period.

@@

The transition to online learning has not gone without hiccups.  Technology and access are two
key factors for success in online learning environments, which were a barrier for many of our
students.  Our institution implemented a number of academic measures intended to support our
students during the pandemic, including extending the withdrawal date and allowing students to
be graded pass/fail after seeing their final grade during each of the Spring 2020 and Fall 2020
semesters. There have also been technical supports put in place at the institution such as
extended laptop loan programs and remote proctoring was paid for by a grant through the
university system. Additionally, our institution's Center for Survey Research published
some findings in August 2020 that explain some of our observed
issues.\citep{ECUcovidSurveyRedacted} The GTAs and students found communication about proposals
much more difficult online than in face-to-face classes.  About \SI{10}{\percent} to
\SI{20}{\percent} of our students did not have reliable internet access after the transition to
online learning in the spring.\citep{ECUcovidSurveyRedacted} Also, up to \SI{8}{\percent} of
students did not have access to a computer once they left campus.\citep{ECUcovidSurveyRedacted}
Additionally, we noticed that some students in the course did not have access to computers
capable of running \emph{Tracker} or \emph{ImageJ}, both of which run on Windows, Macintosh, or
Linux computers but not Chromebooks or mobile devices. We discovered
\emph{jsTrack}\citep{jstrack}, an online Javascript web application for video analysis that
runs on most computers including Chromebooks, but not mobile devices.

>>>>

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and
techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging students
in and assessing scientific practices, we secured funding, both internally and from the NSF, to
change this \emph{status quo} in physics, biology, and chemistry.  All three departments
transformed their labs using the Argument Driven Inquiry
(ADI)\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has been
specifically designed to engage students in authentic science practices. ADI is very much in
the spirit of transformations such as Modeling Physics\cite{modelingBrewe2008} and the
Investigative Science Learning Environment (ISLE) \cite{etkina2007investigative}. Oftentimes,
both Modeling Physics and ISLE are implemented in studio-style or other active/interactive
lecture classrooms, whereas ADI is lab-focused. The common use of hybrid courses for Modeling
and ISLE imply a shorter time scale for their learning cycles, for example, an ISLE cycle often
takes place over the course of a week, whereas the cycles that we use in ADI take place over 4
weeks. For the institutional transformation of our physics labs with ADI, we also utilized the
AAPT Lab Recommendations to inform our curricular choices.\citep{kozminski2014aapt}

@@

We developed lab kits with supplies that allowed students to perform the investigations outside
the teaching laboratory so we could provide students with a hands-on experience. We purchased
the lab kit items in collaboration with our campus bookstore, and the students purchased the
lab kits from the bookstore. We ordered lab kit items in bulk and, where possible, directly from
manufacturers to reduce the costs of the items. Each Physics 1 kit cost \$20.60 and each
Physics 2 kit cost \$39.00. Students paid for these kits in place of a lab manual, which we
provided online for this course. Lab kit materials were used for each investigation. The two video 
analysis investigations in the Physics 1 laboratory (one comparing the accelerations of a marble 
rolling up and down an incline and the other analyzing a marble collision in two dimensions) 
required the students to record their own videos of marbles from their lab kits using their 
smartphone cameras. The image analysis investigation in the 
Physics 2 laboratory required students to cut their own slits in aluminum foil and also create mounts 
for hairs to produce diffraction patterns with lab-kit lasers. The online adaptation of the face-to-face 
ADI curriculum has changed slightly. Pre-labs are done in groups asynchronously.  Proposals are 
developed in groups during the lab period, and scheduled argumentation sessions occur via 
WebEx during the lab period.

@@

The transition to online learning has not gone without hiccups.  Technology and access are two
key factors for success in online learning environments, which were a barrier for many of our
students.  Our institution implemented a number of academic measures intended to support our
students during the pandemic, including extending the withdrawal date and allowing students to
be graded pass/fail after seeing their final grade during each of the Spring 2020 and Fall 2020
semesters. Other supports put in place at the institution included an extended laptop loan
program. Also, remote proctoring was paid for by a grant through the university system. Our
institution's Center for Survey Research published some findings in August 2020 that explain
some of our observed issues.\citep{ECUcovidSurveyRedacted} The GTAs and students found
communication about proposals much more difficult online than in face-to-face classes.  About
\SI{10}{\percent} to \SI{20}{\percent} of our students did not have reliable internet access
after the transition to online learning in the spring.\citep{ECUcovidSurveyRedacted} Also, up
to \SI{8}{\percent} of students did not have access to a computer once they left
campus.\citep{ECUcovidSurveyRedacted} Additionally, we noticed that some students in the course
did not have access to computers capable of running \emph{Tracker} or \emph{ImageJ}, both of
which run on Windows, Macintosh, or Linux computers but not Chromebooks or mobile devices. We
discovered \emph{jsTrack}\citep{jstrack}, an online Javascript web application for video
analysis that runs on most computers including Chromebooks, but not mobile devices.
